---
title: "FinalExam158"
author: "Henrique"
date: "May 6, 2019"
output:
  word_document: default
  html_document: default
---
1)
```{r}
n <- 30
coefs_age <- rep(0,1e4)
coefs_age2 <- rep(0,1e4)
for (i in 1:1e4) {
 Age=round(runif(n,min=18,max=70))
 Age2 <- Age^2
 HR <- 94-Age*0.5+Age2*0.0035+rnorm(n,sd=10)
 model <- lm(HR~Age+Age2)
 coefs_age[i] <- summary(model)$coefficients[2,1]
 coefs_age2[i] <- summary(model)$coefficients[3,1]

}
mean(coefs_age)
mean(coefs_age2)
summary(model)


```
a)mean of coefs_age is -0.50 and coefs_age2 is 0.004, its seems right as the summary of one of the instances of the model shows as coef for age being -.53 and for age2 to be -.005 which is pretty close to the mean.

b)
```{r}
n <- 30
ageee <- rep(0,1e4)
HRrr <- rep(0,1e4)
set.seed(104)
for (i in 1:1e4) {
 Age=round(runif(n,min=18,max=70))
 Age2 <- Age^2
 HR <- 94-Age*0.5+Age2*0.0035+rnorm(n,sd=10)
 model <- lm(HR~Age+Age2)
#ageee[i]<- mean(Age)
#HRrr[i]<- mean(HR)

}
plot(Age,HR, xlab='Age', ylab = 'Heart Rate', main = 'Heart Rate vs Age', cex=0.9)

```
c)
```{r}
n <- 30
p1 <- rep(0,1e4)
p2 <- rep(0,1e4)
for (i in 1:1e4) {
 Age=round(runif(n,min=18,max=70))
 Age2 <- Age^2
 HR <- 94-Age*0.5+Age2*0.0035+rnorm(n,sd=10)
 model <- lm(HR~Age+Age2)
 p1[i] <- summary(model)$coefficients[2,4]
 p2[i] <- summary(model)$coefficients[3,4]
}
sum(p1<0.05)/length(p1)
sum(p2<0.05)/length(p2)

```
Power of age is 0.09  and the power of age^2 is 0.06

d)
```{r}
n <- 30
p1 <- rep(0,1e4)
for (i in 1:1e4) {
 Age=round(runif(n,min=18,max=70))
 Age2 <- Age^2
 HR <- 94-Age*0.5+Age*0.0035+rnorm(n,sd=10)
 model <- lm(HR~Age)
 p1[i] <- summary(model)$coefficients[2,4]
}
sum(p1<0.05)/length(p1)
summary(model)

```
It increased the power by a lot which would mean that the relation between HR and age it is not quadratic.




2)a)
```{r}

h1 <- rgamma(10000, shape = 0.5)
h2 <- rgamma(10000, shape =0.1)
h3 <- rgamma(10000, shape =0.05)

hist(h1,breaks=100)
hist(h2,breaks=100)
hist(h3,breaks=100)



```
As shape gets smaller the histogram gets more skewed to the right.

b)
```{r}

tstat <- rep(1e4)
pvav <- rep(1e4)

for (i in 1:1e4) {
sim <- rgamma(30, shape = 1)
sim1 <- rgamma(30, shape = 1)

t <- t.test(sim,sim1)
tstat[i] <- t$statistic
pvav[i] <- t$p.value
}
hist(tstat)
sum(pvav<0.05)/length(pvav)
  
```
c)
```{r}
tstat <- rep(1e4)
pvav <- rep(1e4)

for (i in 1:1e4) {
sim <- rgamma(30, shape = 0.5)
sim1 <- rgamma(30, shape = 0.5)

t <- t.test(sim,sim1)
tstat[i] <- t$statistic
pvav[i] <- t$p.value
}
hist(tstat)
sum(pvav<0.05)/length(pvav)

tstat <- rep(1e4)
pvav <- rep(1e4)

for (i in 1:1e4) {
sim <- rgamma(30, shape = 0.1)
sim1 <- rgamma(30, shape = 0.1)

t <- t.test(sim,sim1)
tstat[i] <- t$statistic
pvav[i] <- t$p.value
}
hist(tstat)
sum(pvav<0.05)/length(pvav)
```
the distribution is less normal as the shape decreases, also as the shape gets smaller the proportion of pvalues under 0.05 also gets smaller.


```{r}
tstat <- rep(1e4)
pvav <- rep(1e4)

for (i in 1:1e4) {
sim <- rgamma(10, shape = 1)
sim1 <- rgamma(10, shape = 1)

t <- t.test(sim,sim1)
tstat[i] <- t$statistic
pvav[i] <- t$p.value
}
hist(tstat)
sum(pvav<0.05)/length(pvav)
tstat <- rep(1e4)
pvav <- rep(1e4)

for (i in 1:1e4) {
sim <- rgamma(10, shape = 0.5)
sim1 <- rgamma(10, shape = 0.5)

t <- t.test(sim,sim1)
tstat[i] <- t$statistic
pvav[i] <- t$p.value
}
hist(tstat)
sum(pvav<0.05)/length(pvav)
tstat <- rep(1e4)
pvav <- rep(1e4)

for (i in 1:1e4) {
sim <- rgamma(10, shape = 0.1)
sim1 <- rgamma(10, shape = 0.1)

t <- t.test(sim,sim1)
tstat[i] <- t$statistic
pvav[i] <- t$p.value
}
hist(tstat)
sum(pvav<0.05)/length(pvav)
```
yes, as the shape and the n go down the distribution of the mean is less and less normal.






3a)
```{r}
ferret = read.csv("Ferret_Vaccine.csv", header = TRUE)
attach(ferret)
mean(Temperature)
sd(Temperature)

mean(Weight)
sd(Weight)

detach(ferret)

```
Temperature mean = 101.7395
Temperature SD= 0.9505654
Weight mean = 1145.234
Weight SD = 148.4198


b)
```{r}
attach(ferret)

plot(Day[Ferret.ID==574],Temperature[Ferret.ID==574],type="l",xlab='Day',ylab = 'Temperature', ylim=c(100,105))
lines(Day[Ferret.ID==546],Temperature[Ferret.ID==546],lty = 2)
lines(Day[Ferret.ID==548],Temperature[Ferret.ID==548], lty = 3)

detach(ferret)
```

b)
```{r}

```






