---
title: "HWK5"
author: "Henrique"
date: "April 14, 2019"
output: word_document
---



Question 1a)
```{r}


jan = read.csv("january_temp.csv", header = TRUE)

attach(jan)
plot(Lat,Long,xlab= "Latitude",ylab= "Longitude")


detach(jan)


```
Yes, it looks scattered as I expected.


Question 1b)
```{r}

attach(jan)
estjan <- lm(JanTemp~Lat)
summary(estjan)
plot(JanTemp~Lat)

objects(estjan)
abline(estjan$coefficients)


JanTemp=108.73-2.11*Lat

detach(jan)




```

The R squared statistisc is 0.7192.


Question 1c)


```{r}
attach(jan)
estjan2 <- lm(JanTemp~Long)
summary(estjan2)
plot(JanTemp~Long)

objects(estjan2)
abline(estjan2$coefficients)


JanTemp2=1.97-2.11*Long

detach(jan)


```
the R squared statistics is 0.02096.



question 1)d)
latitude had a much better prediction, which is shown by its r squared statistic being 1.




Question 1)e)

```{r}

attach(jan)
estjan3 <- lm(JanTemp~Long+Lat)
summary(estjan3)



JanTemp3=108.73-2.11*Lat*Long

detach(jan)

```
Lat was much better at predicting. lat coefficient changed but long didn't 

question 1)f)
```{r}




```


question 1)g)

```{r}

  anova(estjan,estjan2,estjan3)


```



question 2)


```{r}
x=rnorm(100,10,10)
z = t.test(x)
z$statistic
lol <- rep(0,2)
lol[1]<-z$statistic
```


```{r}



t1t <- rep(0,10e4)
for (i in 1:length(t1t)){
  x=rnorm(100,10,10)
  z = t.test(x)
  t1t[i] <- z$statistic
}

hist(t1t)

```




question 3a)
```{r}

t2t <- rep(0,10000)
for (i in 1:10000){
  x<-(rnorm(100,5,1)+rnorm(10,5,5))
  z = t.test(x)
  t2t[i] <- (z$statistic)
  
}


hist(t2t)



```

3b)
```{r}
t2t <- rep(0,10000)
for (i in 1:10000){
  x<-(rnorm(100,5,1)+rnorm(10,5,5))
  z = t.test(x,var.equal = TRUE)
  t2t[i] <- (z$statistic)
  
}


hist(t2t)
```

part c looks less skewed to the right when compared to part bv

















