---
title: 'Homework #11'
author: "Henrique Magalhaes Rio"
date: "11/19/2019"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
# Insert packages you need here
library(knitr)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(broom)
library(splines)

```

# Question 1 

```{r, include=FALSE}

words <-read.csv("english_words.csv")

```

## Part 1 (a)

```{r, echo=FALSE}

ggplot(words,aes(y=WrittenFrequency,x=LengthInLetters)) +theme_bw()+ geom_point() +geom_smooth(method=lm, se=FALSE)+ggtitle("frenquency of words vs lenght of the words")+xlab("Lenght of the word")+ylab("Frenquency of the word")
wordslm <- lm(WrittenFrequency~LengthInLetters, data =words)


```
## Part 2 (b)


```{r, echo=FALSE}

ggplot(words,aes(y=log(WrittenFrequency),x=LengthInLetters)) +theme_bw()+ geom_point() +ggtitle("log of frenquency of words vs lenght of the words")+xlab("Lenght of the word")+ylab("log(Frenquency of the word)")



```



## Part 3 (c)

```{r,include=FALSE}

wordslm2 <- lm(log(WrittenFrequency)~LengthInLetters, data =words)

```
$\hat\beta_0$=5.65479
$\hat\beta_1$=-0.146


## Part 4 (d)

A difference of 1 percentage point in the lenght of the words is associated with a negative difference of 0.146 (95% CI: -0.236,-0.0562) in the estimated average  of the logarithm of word frenquency.

## Part 4 (e)

A difference of 1 percentage point in the lenght of the words is associated with a multiplicative  difference of -0.146 (95% CI: -0.236,-0.0562) in the geometric mean of word frequency.



# Question 2

```{r, include=FALSE}
psa <-read.csv("psa_complete.csv")

```

## Part 1 (a)

```{r,echo=FALSE}

ggplot(psa,aes(y=obstime,x=nadirpsa)) +theme_bw()+ geom_point() + ylab("Days in remission")+ xlab("lowest concentration of PSA")

```



## Part 2 (b)

```{r,echo=FALSE}

ggplot(psa,aes(y=obstime,x=log(nadirpsa))) +theme_bw()+ geom_point()+ ylab("Days in remission")+ xlab("log  of lowest concentration of PSA")

```

## Part 3 (c)

```{r,include=FALSE}

psalm <- lm(obstime~log(nadirpsa),data=psa)

```

-3.93 is the difference in time spent in remession after treatment for a e-fold difference in the lowest concentration of PSA


# Question 3 

## Part 1 (a)
```{r,include=FALSE}

subject1 <- subset(Theoph, Subject==1)
```

## Part 2 (b)

```{r, echo = FALSE}
ggplot(subject1,aes(y=conc,x=Time)) +theme_bw()+ geom_point()+ylab("Concentration")
```


## Part 3 (c)

```{r, echo = FALSE}
ggplot(subject1,aes(y=conc,x=Time)) +theme_bw()+ geom_point() +ylab("Concentration")+ geom_smooth(aes(x=Time, y=conc),
data=subject1,
method="lm",
formula=y~ns(x, df=3))


```

## Part 4 (d)

```{r, include=FALSE}
sublm <- lm(conc~ns(Time,3), data = subject1)
```

$\hat\beta_0$=3.10
$\hat\beta_1$=-3.29
$\hat\beta_2$=7.97
$\hat\beta_3$=-3.94

## Part 5 (e)

```{r, echo = FALSE}
ggplot(subject1,aes(y=conc,x=Time)) +theme_bw()+ylab("Concentration") +geom_point() + geom_smooth(aes(x=Time, y=conc),
data=subject1,
method="lm",
formula=y~ns(x, df=4))


```

## Part 6 (f)

```{r,include=FALSE}
sublm2 <- lm(conc~ns(Time,4), data = subject1)
```

$\hat\beta_0$=1.28
$\hat\beta_1$=5.86
$\hat\beta_2$=0.821
$\hat\beta_3$=14.1
$\hat\beta_4$=-3.93

## Part 7 (g)
based on the $R^2$ the model with 4 splines ($R^2$=0.927) is a lot better then the model with 3 splines ($R^2$=0.6895)



# Appendix

```{r show-code, ref.label = all_labels(), echo = TRUE, eval = FALSE}

```