---
title: 'Homework #2'
author: "Henrique Magalhaes Rio"
date: "9/4/2019"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
# Insert packages you need here
library(knitr)
library(ggplot2)
library(dplyr)
library(tidyverse)

```


# Question 1

Designed studies are much better at predicting as they are not subject to bias of an observational studie.



# Question 2

## Part 1 (a)
it comes from cars from the years 19

## Part 2 (b)
```{r, echo=FALSE, include=FALSE}
summarize(mpg, mean(hwy))
```
the mean on the highway is 23.4 mpg

## Part 3 (c)
```{r, include=FALSE, echo=FALSE}


mpg %>%
    group_by(drv) %>%
    summarize(n=n()) %>%
    mutate(freq = n / sum(n))



```

there are 103 4 wheel drive vehicles, 106 front wheel drive vehicles and 25 rear wheel drive vehicles.


## Part 4 (d)

```{r, include=FALSE,echo=FALSE}

mpg %>%
  group_by(drv) %>%
  summarize(mean_mpg=mean(hwy))
  

```
The mean for 4 wheel drive cars is 19.17, for front wheel drive 28.16038 and rear wheel drive 21.000.

# Question 3

## Part 1 (a)
```{r, echo=FALSE}
ggplot() + geom_point(aes(x=displ,y=hwy, col=drv),data=mpg) + xlab("Engine Displacement") + ylab("Highway miles per gallon")
```
  
## Part 2 (b)
  
```{r, echo=FALSE}
ggplot() + geom_point(aes(x=displ,y=hwy, col=class),data=mpg) + xlab("Engine Displacement") + ylab("Highway miles per gallon")
```

## Part 3 (c)

```{r, echo=FALSE}
ggplot() + geom_point(aes(x=displ,y=hwy, col=class),data=mpg) + xlab("Engine Displacement") + ylab("Highway miles per gallon")+xlim(c(0,10))
```

# Question 4

## Part 1 (a)
```{r, echo=FALSE}
ggplot(mpg) + geom_histogram(aes(x=cty))+xlab("City miles per gallon")

```

## Part 2 (b)
```{r, echo= FALSE}
ggplot(mpg) + geom_density(aes(x=cty))+xlab("City miles per gallon")

```

## Part 3 (c)
```{r, echo= FALSE}
ggplot(mpg) + geom_density(aes(x=cty,y=stat(count)))+xlab("City miles per gallon")

```
## Part 4 (d)
```{r, echo= FALSE}
ggplot(mpg) + geom_density(aes(x=cty,y=stat(count)))+geom_histogram(aes(x=cty))+xlab("City miles per gallon")

```

# Question 5

## Part 1 (a)
$E[3X]= 15$

## Part 2 (b)
$E[2X + 4Y -5] = 21$

## Part 3 (c)
$Var(2Y-3)= 4$

## Part 4 (d)
$Var[2X + 4Y] = 20$

# Question 6

$Y_i = \beta_0 + \beta_1x_i$


# Appendix

```{r show-code, ref.label = all_labels(), echo = TRUE, eval = FALSE}

```