---
title: 'Homework ## Template'
author: "Your Name Here"
date: "Date Here"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
# Insert packages you need here
library(knitr)
library(ggplot2)

```


# About this template

This file provides a template that you should use for homework assignments. 

## R Markdown 

Markdown is a simple formatting syntax for writing HTML, PDF, and Word documents using a common interface. This template is setup to generate pdf or MS Word (doc) files, since those are well-supported by the Canvas system. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

You can enter text into the document using the editor built into RStudio. When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r iris, echo=TRUE}
# R code goes in blocks like this one.
# Use comments to help document your code
head(iris)
```

In general, your homework responses should not include R code or raw R output unless the question specifically requests them.  To prevent them from being printed, set `include=FALSE` in the code block header. For example:

```{r iris2, include=FALSE}
# Viewing the iris dataset, the second time
head(iris)
```

Documentation for other chunk options is available at <https://yihui.name/knitr/options/>

## Headings

If using this template for your homework, delete the text within these sections and add your homework answers.

Type `# Question (number here)` on a line to create a header for each question. For problems with sub-parts, a secondary heading can be created by typing `## Part (letter here)` on a new line.

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE,  out.width="75%"}
# Code for the plot here
ggplot(pressure) + 
    geom_point(aes(x=temperature, y=pressure)) +
    xlab("Temperature (deg C)") + ylab("Pressure (mmHg)")
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot. It would not work in this case to set `include=FALSE`, since that would also hide the plot output.

## Including Math

RMarkdown allows you to easily incorporate formatted mathematical notation using LaTeX. Mathematical expressions should be typed between dollar signs (\$). For example, we can write $x + y$. 

* Subscripts can be added using underscore: $x_1$
* Superscripts can be added using the caret character: $x^2$
* Greek letters can be added by typing \\ followed by the letter name: $\beta$
* Equations can be centered and set off from text using two dollar signs: $$a^2 + b^2 = c^2$$


Note: Creating a MS Word document does not require an installation of LaTeX on your computer. Creating a PDF output file does. A simple way to do this to install 'tinytex' by running the following commands:

```{r echo=TRUE, eval=FALSE}
# Only need to do this *once* for your computer
# in order to be able to create PDFs with math notation
install.packages('tinytex')
tinytex::install_tinytex()
```



# Question 1

## Part 1(a)
Responses here

## Part 1(b)
Responses here

# Question 2
Responses here.

# Question 3

The figure for this question is:

```{r pressure2, echo=FALSE,  out.width="75%"}
# Code for the plot here
ggplot(pressure) + 
    geom_point(aes(x=temperature, y=pressure)) +
    xlab("Temperature (deg C)") + ylab("Pressure (mmHg)")
```


# Appendix

```{r show-code, ref.label = all_labels(), echo = TRUE, eval = FALSE}

```